---
title: 'STATS 205: Homework Assignment 4'
author: "Brian Liu"
date: "6/3/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution to Problem 1

```{r}
allergics = c(1651.0, 1112.0, 102.4, 100.0, 67.6, 65.9, 64.7, 39.6, 31.0)
nonallergics = c(48.1, 48.0, 45.5, 41.7, 35.4, 34.3, 32.4, 29.1, 27.3, 18.9, 6.6, 5.2, 4.7)
allergics; nonallergics
```

**Assumption**: The two samples are from independent populations.

>Two data samples are independent if they come from distinct populations and the samples do not affect each other.
>-- <cite>[Mann-Whitney-Wilcoxon Test][1]</cite>

[1]:http://www.r-tutor.com/elementary-statistics/non-parametric-methods/mann-whitney-wilcoxon-test

$H_0$: Allergic smokers have the sputum histamine levels as nonallergic smokers. That is,

\[
    H_0: p_a = p_n
\]

$H_A$: Allergic smokers have higher sputum histamine levels than nonallergic smokers. That is,

\[
    H_0: p_a > p_n
\]

