---
title: 'STATS 205: Homework Assignment 4'
author: "Brian Liu"
date: "6/3/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution to Problem 1

```{r}
allergics = c(1651.0, 1112.0, 102.4, 100.0, 67.6, 65.9, 64.7, 39.6, 31.0)
nonallergics = c(48.1, 48.0, 45.5, 41.7, 35.4, 34.3, 32.4, 29.1, 27.3, 18.9, 6.6, 5.2, 4.7)
allergics; nonallergics
```

The null hypothesis is that allergic smokers have the same sputum histamine levels as nonallergic smokers. That is,

\[
    H_0: p_a = p_n
\]

The alternative hypothesis is that allergic smokers have higher sputum histamine levels than nonallergic smokers. That is,

\[
    H_0: p_a > p_n
\]

To test the null hypothesis against the alternative hypothesis, we will use the Mann-Whitney-Wilcoxin test, since the two samples are independent.

>Two data samples are independent if they come from distinct populations and the samples do not affect each other.
>-- <cite>[Mann-Whitney-Wilcoxon Test][1]</cite>

[1]:http://www.r-tutor.com/elementary-statistics/non-parametric-methods/mann-whitney-wilcoxon-test

```{r}
wilcox.test(x = allergics, y = nonallergics, alternative = "greater")
```

The $p$-value is $0.000386$, which is significant at the $\alpha = 0.05$ level. There is strong evidence that allergic smokers have higher sputum histamine levels than nonallergic smokers.

## Solution to Problem 2

[Original problem statement](https://i.imgur.com/WXQHBJ9.png)

```{r}
karate = c(37, 39, 30, 7, 13, 139, 45, 25, 16, 146, 94, 16, 23, 1, 290, 169, 62, 145, 36, 20, 13)
olympics = c(12, 44, 34, 14, 9, 19, 156, 23, 13, 11, 47, 26, 14, 33, 15, 62, 5, 8, 0, 154, 146)
karate; olympics
```

The null hypothesis is that children who viewed the violent TV take the same amount of time to seek help (were as tolerant) as the children who viewed the nonviolent sports-action TV. That is, 

\[
    H_0: t_k = t_o
\]

The alternative hypothesis is that children who viewed the violent TV take longer to seek help (were more tolerant) than the children who viewed the nonviolent sports-action TV. That is,

\[
    H_0: t_k > t_o
\]

```{r}
wilcox.test(x = karate, y = olympics, alternative = "greater")
```

The $p$-value is $0.08126$, which is *not* significant at the $\alpha = 0.05$ level. There is *not enough* evidence that children who viewed the violent TV take longer to seek help (were more tolerant) than the children who viewed the nonviolent sports-action TV.

## Solution to Problem 3

Let $X$ be the nonallergics and $Y$ be the allergics.

\[
    \delta = P(X < Y)
\]


```{r}
allergics = c(1651.0, 1112.0, 102.4, 100.0, 67.6, 65.9, 64.7, 39.6, 31.0)
nonallergics = c(48.1, 48.0, 45.5, 41.7, 35.4, 34.3, 32.4, 29.1, 27.3, 18.9, 6.6, 5.2, 4.7)
allergics; nonallergics

wilcox.test(x = allergics, y = nonallergics, conf.int=TRUE, conf.level=.90)
```

<!-- The $p$-value stands for how likely the sample data would occur, assuming that the null hypothesis (that allergic smokers have the same sputum histamine levels as nonallergic smokers, or $H_0: p_a = p_n$) is true.
 -->
The estimate for $\delta$ is

\[
    \hat{\delta} = P(X < Y) = 54.3
\]

and the $90%$ confidence interval for $\delta$ is

\[
    \hat{\delta} = P(X < Y) = (25.9, 81.1)
\]


## Solution to Problem 4

```{r}
term = c(0.80, 0.83, 1.89, 1.04, 1.45, 1.38, 1.91, 1.64, 0.73, 1.46)
gest = c(1.15, 0.88, 0.90, 0.74, 1.21)

ks.test(x = term, y = gest, alternative = "greater")
```

