---
title: 'STATS 205: Homework Assignment 3'
author: "Brian Liu"
date: "4/26/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution to Problem 1

### (i)

```{r}
library(bootstrap); data(law)
t(law)
theta.hat = cor(law$LSAT, law$GPA); theta.hat
library(partitions)
n = 15
allCompositions = compositions(n, n);allCompositions[,1:5]
allCompositions.sub = allCompositions[, sample(1:dim(allCompositions)[2], size=10000, replace=FALSE)]

draw.bootstrap.samples = function(df){
  n = dim(df)[1]
  ind = sample(n, replace = TRUE)
  cor.bootstrap.replicate = cor(df[ind, "LSAT"], df[ind, "GPA"])
  return(cor.bootstrap.replicate)
}
R = 10000
theta.hat.star = replicate(R, draw.bootstrap.samples(law))
# make a ggplot
library(ggplot2)
theta.hat.star.df = data.frame(theta.hat.star = theta.hat.star)
ggplot(theta.hat.star.df) +
  geom_density(aes(x = theta.hat.star, y = ..scaled..),
    fill = "lightblue") +
  geom_hline(yintercept=0, colour="white", size=1) +
  theme_bw() +
  ylab("density") +
  xlab(bquote(hat(theta))) +
  geom_vline(xintercept = theta.hat, col = "red")+
  scale_y_continuous(expand = c(0,0))
```

### (ii)

```{r}
sd(theta.hat.star)
```

## Solution to Problem 2

### (i)

```
67 runs resulting in swallowing attempts
58 successful
9 failed

H_0 : p = 0.6
H_A : p > 0.6


```

```{r}
n = 67
successes = 58
pbar = successes / n; pbar

p0 = 0.6; nsim = 10000
B = rbinom(nsim, size = n, prob = p0)
hist(B, breaks = 30)
```

Test statistic $Z$:

\[
    Z_0 = \frac{B - 67(0.6)}{(67(0.6)(0.4))^{\frac{1}{2}}}
\]

```{r}
qnorm((1-0.05), mean = 0, sd = 1)
```

Rejection region: $Z \geq z_{0.05} = 1.645$

Observed test statistic $Z_o$:

\[
    Z_o = \frac{58 - 67(0.6)}{(67(0.6)(0.4))^{\frac{1}{2}}} = 4.44
\]

```{r}
numerator = successes - (n * p0)
denominator = sqrt(n * p0 * (1.0 - p0))
Z.obs = numerator / denominator; Z.obs
```

The large sample approximation value $Z_o = 2.5 > 1.645$ and thus we reject $H_0 : p = 0.6$ in favor of $p > 0.6$ at the approximate $\alpha = 0.05$ level. Thus there is evidence that the success rate of swallowing attempts is greater than $0.6$.

### (ii)

Power is the probability of rejecting $H_0$ when $H_A$ is true. We found that test reject $H_0$ is $Z \geq z_{0.05} = 1.645$. Therefore, if $p = 0.7$,

\[
    Z_o = \frac{58 - 67(0.6)}{(67(0.6)(0.4))^{\frac{1}{2}}} = 4.44
\]

is no longer standard normal.

We have 

\[
    Z_{o7} = \frac{58 - 67(0.7)}{(67(0.7)(0.3))^{\frac{1}{2}}} = 2.96
\]

```{r}
p1 = 0.7
numerator = successes - (n * p1)
denominator = sqrt(n * p1 * (1.0 - p1))
Z.obs.seven = numerator / denominator; Z.obs.seven
```

\[
    Power = P(Z \geq 1.645 | p = 0.7)
\]

\[
    = P_{p = 0.7}\bigg(\frac{B - 67(0.6)}{(67(0.6)(0.4))^{\frac{1}{2}}} \geq 1.645\bigg)
\]

\[
    = P_{p = 0.7}(B \geq 1.645(67(0.6)(0.4))^{\frac{1}{2}} + 67(0.6))
\]

\[
    = P_{p=0.7}\bigg(\frac{B - 67(0.7)}{(67(0.7)(0.3))^{\frac{1}{2}}} \geq \frac{1.645(67(0.6)(0.4))^{\frac{1}{2}} + 67(0.6) - 67(0.7)}{(67(0.7)(0.3))^{\frac{1}{2}}}\bigg)
\]


```{r}
triple_product = n * p0 * (1.0 - p0)
first_term = 1.645 * sqrt(triple_product)
second_term = n * p0
third_term = n * p1
bottom_term = n * p1 * (1.0 - p1)
p7_numerator = first_term + second_term - third_term
p7_denominator = sqrt(bottom_term)
Pp_7_zvalue = p7_numerator / p7_denominator; Pp_7_zvalue
```

\[
    P(Z^* \geq -0.0276) = 0.4890
\]

```{r}
pvalue = pnorm(Pp_7_zvalue); pvalue
```

If $p = 0.8$,

\[
    Power = P(Z \geq 1.645 | p = 0.8)
\]

\[
    = P_{p=0.8}\bigg(\frac{B - 67(0.8)}{(67(0.8)(0.2))^{\frac{1}{2}}} \geq \frac{1.645(67(0.6)(0.4))^{\frac{1}{2}} + 67(0.6) - 67(0.8)}{(67(0.8)(0.2))^{\frac{1}{2}}}\bigg)
\]


```{r}
p2 = 0.8
triple_product = n * p0 * (1.0 - p0)
first_term = 1.645 * sqrt(triple_product)
second_term = n * p0
third_term = n * p2
bottom_term = n * p2 * (1.0 - p2)
p8_numerator = first_term + second_term - third_term
p8_denominator = sqrt(bottom_term)
Pp_8_zvalue = p8_numerator / p8_denominator; Pp_8_zvalue
```

\[
    P(Z^* \geq -2.078) = 0.01886
\]

```{r}
# pvalue = pnorm(-abs(Pp_7_zvalue)); pvalue
pvalue = pnorm(Pp_8_zvalue); pvalue
```


<!-- There is a function in R that calculates the same thing: library(pwr), pwr.p.test()-->

## Solution to Problem 3

**Summary**: **Estimate for $\hat{p} = 0.8615$** and **estimate for standard deviation of $\hat{p} = 0.04284$**.

Estimate for $p$ using binomial confidence interval, `binom.confint()`:

```{r}
library(binom)
binom.confint(x=56, n=65, conf.level=.95, methods = "asymptotic")
```

\[
    \hat{p} = (0.7776, 0.9455)
\]

Estimate for $p$ using 1-sample proportions test without continuity correction, `prop.test()`:

```{r}
prop.test(x=56, n=65, p = 0.6, conf.level=0.95, alternative = c("greater"))
```

\[
    p = 0.8615
\]

Estimate for $p$ using Exact Binomial Test:

```{r}
binom.test(x=56, n=65, p = 0.6, alternative = c("greater"), conf.level = 0.95)
```

Standard error of $\hat{p}$ is:

\[
    \sqrt{\frac{p(1 - p)}{n}}
\]
\[
    = \sqrt{\frac{(0.6)(0.4)}{65}}
\]
\[
    = 0.06076
\]

```{r}
p = 0.6
n = 65
numerator = p * (1 - p)
denominator = n
answer = sqrt(numerator/denominator); answer
```

and estimate is:

\[
    \sqrt{\frac{\hat{p}(1 - \hat{p})}{n}}
\]
\[
    = \sqrt{\frac{(0.8615)(1 - 0.8615)}{65}}
\]
\[
    = 0.04284
\]

```{r}
p.hat = 0.8615
n = 65
numerator = p.hat * (1 - p.hat)
denominator = n
answer = sqrt(numerator/denominator); answer
```

## Solution to Problem 4

```{r}

binom.confint(x = 56, n = 65, conf.level = 0.96, methods = "all")
```

Here are the rows relevant to our problem:

```
##           method  x  n      mean     lower     upper
## 1  agresti-coull 56 65 0.8615385 0.7488973 0.9301180
## 2     asymptotic 56 65 0.8615385 0.7735567 0.9495202
## 5          exact 56 65 0.8615385 0.7480632 0.9371740
## 11        wilson 56 65 0.8615385 0.7514483 0.9275670
```

where `asymptotic` is Laplace-Wald, `agresti-coull` is Agresti-Coull, `exact` is Clopper-Pearson, and `wilson` is Wilson.

It looks like `agresti-coull` and `exact` are fairly similar in terms of the location of the interval, while `asymptotic` is skewed towards the "right" side of the intervals, and `wilson` has the smallest range.

## Solution to Problem 5

```{r}
plant_vector <- c(926, 288, 293, 104)

expected = c(9/16, 3/16, 3/16, 1/16)

# goodness-of-fit test

res <- chisq.test(x = plant_vector, p = expected); res

res$expected
res$observed
```

The **p-value** of the test is $p = 0.6895$, which is more than the significance level $\alpha = 0.05$. We cannot conclude that the data does *not* support Mendelian theory at the $\alpha = 0.05$ level.

## Solution to Problem 6

```{r}
df = data.frame(Low = c(0, 2), High = c(5, 1))
rownames(df) = c("Multiple attack", "Primary attack"); df
fisher.test(df, alternative = "less")
```

There is *not enough evidence* that the unknown probability that a multiple-attack patient will have low reactivity is *less than* the unknown probability that a primary-attack patient will have low reactivity at the $\alpha = 0.05$ level. The $P$-value achieved by these data if we use Fisher's exact test of $H_0$ against the alternative $p_1 < p_2$ is $0.1071$.

```{r}
df = data.frame(Low = c(0, 2), High = c(5, 1))
rownames(df) = c("Multiple attack", "Primary attack"); df
fisher.test(df, alternative = "two.sided")
```

The $P$-value achieved when testing the null against a two-sided hypothesis is the same, $0.1071$, and therefore, there is *not enough evidence* that the unknown probability that a multiple-attack patient will have low reactivity is *not equal* to the unknown probability that a primary-attack patient will have low reactivity.

## Solution to Problem 7

Since we want to find out if having tonsils reduces the rate of Hodgkin's, we want to see if there is a higher rate of tonsillectomy among Hodgkin's cases than among non-Hodgkin's cases. Therefore, we will test the null hypothesis that there is no difference in the rate of tonsillectomy between the two populations of Hodgkin's and non-Hodgkin's cases against the alternative hypothesis that there is higher rate of tonsillectomy among Hodgkin's cases.

\[
    H_0 : p_1 = p_2
\]

\[
    H_A : p_1 > p_2
\]

```{r}
x = c(67, 43)
n = c(101, 107)
prop.test(x, n, alternative = c("greater"), conf.level = 0.99)
```

The 99% confidence interval for difference between rate of tonsillectomy in Hodgkin's cases and rate of tonsillectomy in non-Hodgkin's cases is between 9.66% and 100%.

The $P$-value achieved by these data if we use a `2-sample test for equality of proportions with continuity correction` is $0.0001379$. There is *strong evidence* at the $\alpha = 0.05$ level that there is higher rate of tonsillectomy among Hodgkin's cases.

## Solution to Problem 8

```{r}
elisa_matrix <- matrix(c(82, 13, 6, 0), nrow=2, ncol=2, byrow=TRUE, dimnames = list("ABC-ELISA" = c("+", "-"), "Standard ELISA" = c("+", "-"))); elisa_matrix


```