---
title: 'STATS 205: Final Project Write-Up'
author: "Brian Liu"
date: "6/14/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Background of the data and why it is interesting or important

>The data we are using is the data from WHO suicide statistics from Kaggle. This gives population-based statistics on suicide rate...

# 2. Explanation of the method studied and its properties

# 3. Data analysis or simulation study

We will use the crude rate of suicide per 100,000 people.

>This analysis provides information on age-standardized rates...

```{r}
who_suicide_statistics_df <- read.csv("who_suicide_statistics.csv")
nrow(who_suicide_statistics_df)
colnames(who_suicide_statistics_df)
```

>Filter and save countries with missing suicide rate.

<!-- install.packages("devtools", dependencies=TRUE, repos='http://cran.us.r-project.org') -->

```{r}
library(tidyverse)
filtered_suicide_df <- drop_na(who_suicide_statistics_df, "suicides_no")
nrow(filtered_suicide_df)
write.csv(filtered_suicide_df, 'filtered_suicide.csv')
```

>After filtering countries with missing suicide rate, take a random sample of 100 countries and make sure each continent has approximately equal countries.

\[
	\frac{100 \textrm{ countries}}{7 \textrm{ continents}} \approx 14 \textrm{ to } 15 \textrm{ countries per continent}
\]

Filter countries by continent:


```{r}
install.packages("countrycode", dependencies=TRUE, repos='http://cran.us.r-project.org')
library(countrycode)
df <- data.frame(country = c("Afghanistan",
                             "Algeria",
                             "USA",
                             "France",
                             "New Zealand",
                             "Fantasyland"))
df$continent <- countrycode(sourcevar = df[, "country"],
                            origin = "country.name",
                            destination = "continent")

# Get seven dataframes, filtered by list of countries for each continent.

```

# 4. Interpretation of the results or discussion
