---
title: 'STATS 205: Final Project Write-Up'
author: "Brian Liu"
date: "6/14/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1. Background of the data and why it is interesting or important

>The data we are using is the data from WHO suicide statistics from Kaggle. This gives population-based statistics on suicide rate...

# 2. Explanation of the method studied and its properties

# 3. Data analysis or simulation study

We will use the crude rate of suicide per 100,000 people.

>This analysis provides information on age-standardized rates...

```{r}
who_suicide_statistics_df <- read.csv("who_suicide_statistics.csv")
head(who_suicide_statistics_df)
colnames(who_suicide_statistics_df)
```

>Filter and save countries with missing suicide rate.

<!-- install.packages("devtools", dependencies=TRUE, repos='http://cran.us.r-project.org') -->

```{r}
library(tidyverse)
filtered_suicide_df <- drop_na(who_suicide_statistics_df, "suicides_no")
head(filtered_suicide_df)
```

>After filtering countries with missing suicide rate, take a random sample of 100 countries and make sure each continent has approximately equal countries.

Filter countries by continent:


```{r}
library(countrycode)
filtered_suicide_df$continent <- countrycode(sourcevar = filtered_suicide_df[, "country"],
                            origin = "country.name",
                            destination = "continent")

head(filtered_suicide_df)
write.csv(filtered_suicide_df, 'filtered_suicide.csv')
```

Let us find out which continents are counted:

```{r}
# Get list of continents
list_of_continents <- unique(filtered_suicide_df$continent); list_of_continents
```

Therefore,

\[
	\frac{100 \textrm{ countries}}{6 \textrm{ continents}} \approx 16 \textrm{ to } 17 \textrm{ countries per continent}
\]

we should randomly sample 17 countries from each continent.


Notably, there are countries that are not on any of the listed continents. Let us see which ones those are:

```{r}
not_in_a_continent = filtered_suicide_df[is.na(filtered_suicide_df$continent),]
write.csv(not_in_a_continent, 'not_in_a_continent.csv')
head(not_in_a_continent)
unique(not_in_a_continent$country)
```

Let us make the choice not to include these countries in the analysis, since there are only two countries.

```{r}
# Take off `NA` from list of continents
list_of_continents <- list_of_continents[-length(list_of_continents)]
list_of_continents
```


We will now create six dataframes, filtered by list of countries for each continent.

<!-- # europe_suicide = filtered_suicide_df[filtered_suicide_df$continent == 'Europe',]
# head(europe_suicide)
# length(europe_suicide$country)
# americas_suicide = filtered_suicide_df[filtered_suicide_df$continent == 'Americas',]
# head(americas_suicide)
# length(americas_suicide$country)
# asia_suicide = filtered_suicide_df[filtered_suicide_df$continent == 'Asia',]
# head(asia_suicide)
# length(asia_suicide$country)
# oceania_suicide = filtered_suicide_df[filtered_suicide_df$continent == 'Oceania',]
# head(oceania_suicide)
# length(oceania_suicide$country) -->
<!-- # africa_suicide = filtered_suicide_df[filtered_suicide_df$continent == 'Africa',]
# head(africa_suicide)
# length(africa_suicide$country) -->

```{r}
# library(rlist)
countries_per_continent <- list()

for (i in seq_along(list_of_continents))
{
	countries_per_continent[[i]] <- filtered_suicide_df[filtered_suicide_df$continent == list_of_continents[i],]
}

length(countries_per_continent)

length(countries_per_continent)

for (i in seq_along(countries_per_continent))
{
	print(head(countries_per_continent[[i]]))
	print(length(countries_per_continent[[i]]))
	cat("\n")
}
```

[This text links to very important information about why a `for` loop doesn't print anything.](https://stackoverflow.com/questions/14987743/head-function-does-not-work-within-for-loop)[^1]

[Link to Pandoc Markdown formatting](https://rmarkdown.rstudio.com/authoring_pandoc_markdown.html%23raw-tex)

[^1]:Basically, `for` loops are functions themselves. R prints out the result of a command automatically, but functions are not inherently a command, and since `for` loops are functions, nothing will be printed. The solution is to have `print(command())` within the `for` loop to get output for your `for` loop. You will never again spend hours trying to find out why a `for` loop doesn't print anything because you're no longer an R newbie.

Randomly sample 17 countries from each continent:

```{r}
list_of_continents
for (i in seq_along(countries_per_continent))
{
	print(list_of_continents[i])
	countries <- unique(countries_per_continent[[i]]$country)
	print(countries)
	print(length(countries))
	cat("\n")
}
```

Since there are only 5 countries in Oceania and 12 countries in Africa, we will use all 5 countries of Oceania and all 12 countries of Africa.

```{r}
samples_of_countries <- list()
num_samples <- 17
for (i in seq_along(countries_per_continent))
{
	countries <- unique(countries_per_continent[[i]]$country)
	current_sample <- list()
	if (length(countries) >= num_samples)
	{
		current_sample <- sample(countries, 17)
	} else {
		current_sample <- sample(countries, length(countries))
	}
	samples_of_countries[[i]] <- current_sample
}
```

Let's see the countries that we will be sampling:

```{r}
for (i in seq_along(samples_of_countries))
{
	print(list_of_continents[i])
	print(samples_of_countries[[i]])
	print(length(samples_of_countries[[i]]))
	cat("\n")
}
```


<!-- # Make singular list of countries
# countries_to_test <- list()
# library(rlist)
# for (i in seq_along(samples_of_countries))
# {
# 	current_samples <- samples_of_countries[[i]]
# 	# print(current_samples)
# 	append(x = countries_to_test, values = current_samples)
# }

# length(countries_to_test)

# for (i in seq_along(countries_to_test))
# {
# 	print(countries_to_test[[i]])
# }

# # Not run: 
# x <- list(a=1,b=2,c=3)
# x
# list.append(x,d=4,e=5)
# list.append(x,d=4,f=c(2,3))
# x
# ## End(Not run) -->

Let's filter the original dataframe only to include countries that we have sampled:

```{r}
for (i in seq_along(samples_of_countries))
{
	# find out a way to access each country name
	# print each country name
	for (j in seq_along(samples_of_countries[[i]]))
	{
		sample <- samples_of_countries[[i]]
		print(sample)
	}
	cat("\n")
}
```


# 4. Interpretation of the results or discussion

