---
title: 'STATS 205: Homework Assignment 5'
author: "Brian Liu"
date: "6/10/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Solution to Problem 1

>We say that two observations $X_1$ and $X_2$ are *independent* of one another with respect to a collection of events $\mathcal{A}$ if
>
>\[
>     Pr\left\{X_1 \in A \textrm{  and  } X_2 \in B\right\} = Pr\left\{X_1 \in A \right\}Pr\left\{X_2 \in B\right\}
>\]
>where A and B are any two not necessarily distinct sets of outcomes belonging to $\mathcal{A}^3$.

-- <cite>[2.2.1 Independent Observations; Permutation, Parametric, and Bootstrap Tests of Hypotheses; Good, Phillip I][1]</cite>

>In deciding whether your own observations are exchangeable and a permutation test applicable, the key question is the one we posed in the very first chapter: Under the null hypothesis of no differences among the various experimental or survey groups, can we exchange the labels on the observations without significantly affecting the results?

-- <cite>[2.2.2 Exchangeable Observations; Permutation, Parametric, and Bootstrap Tests of Hypotheses; Good, Phillip I][1]</cite>

[1]:https://searchworks.stanford.edu/view/12473402

## Solution to Problem 2

```{r}
cysticerci <- c(28.9, 32.8, 12.0, 9.9, 15.0, 38.0, 12.5, 36.5, 8.6, 26.8);cysticerci
worms_reco <- c(1.0, 7.7, 7.3, 7.9, 1.1, 3.5, 18.9, 33.9, 28.6, 25.0); worms_reco
```

The null hypothesis is that the mean weight of introduced cysticerci *has no correlation with* the mean weight of worms recovered. That is,

\[
    H_0: \tau = 0
\]

The alternative hypothesis is that the mean weight of introduced cysticerci is *positively correlated with* the mean weight of worms recovered. That is,

\[
    H_A: \tau > 0
\]

To test the null hypothesis against the alternative hypothesis, we will use the Kendall test, a distribution-free test for independence based on signs.

```{r}
cor.test(x = cysticerci, y = worms_reco, method = "kendall", alt = "greater")
```

The $p$-value is $0.7578$, which is *not* significant at the $\alpha = 0.05$ level. There is *not enough* evidence that the mean weight of introduced cysticerci is *positively correlated with* the mean weight of worms recovered.

## Solution to Problem 3

```{r}
cysticerci <- c(28.9, 32.8, 12.0, 9.9, 15.0, 38.0, 12.5, 36.5, 8.6, 26.8)
worms_reco <- c(1.0, 7.7, 7.3, 7.9, 1.1, 3.5, 18.9, 33.9, 28.6, 25.0)
cor.test(x = cysticerci, y = worms_reco, method = "kendall", alt = "greater")
```

The estimate for $\tau = -0.1555556$.

## Solution to Problem 4

```{r}
brain_weight = c(515, 286, 469, 410, 461, 436, 479, 198, 389, 262, 536); length(brain_weight)
fiber_count  = c(32500, 26800, 11410, 14850, 23640, 23820, 29840, 21830, 24650, 22500, 26000); length(fiber_count)

library(bootstrap)
theta.hat = cor(brain_weight, fiber_count); theta.hat
# library(partitions)
# n = 1000
# allCompositions = compositions(n, n);allCompositions[,1:5]
# allCompositions.sub = allCompositions[, sample(1:dim(allCompositions)[2], size=1000, replace=FALSE)]

# draw.bootstrap.samples = function(df){
#   n = dim(df)[1]
#   ind = sample(n, replace = TRUE)
#   cor.bootstrap.replicate = cor(df[ind, "LSAT"], df[ind, "GPA"])
#   return(cor.bootstrap.replicate)
# }
# R = 1000
# theta.hat.star = replicate(R, draw.bootstrap.samples(law))
# # make a ggplot
# library(ggplot2)
# theta.hat.star.df = data.frame(theta.hat.star = theta.hat.star)
# ggplot(theta.hat.star.df) +
#   geom_density(aes(x = theta.hat.star, y = ..scaled..),
#     fill = "lightblue") +
#   geom_hline(yintercept=0, colour="white", size=1) +
#   theme_bw() +
#   ylab("density") +
#   xlab(bquote(hat(theta))) +
#   geom_vline(xintercept = theta.hat, col = "red")+
#   scale_y_continuous(expand = c(0,0))
```


<!-- 
\[
    \hat{\tau} = \frac{2K}{n(n - 1)} = \bar{K}
\]
 -->
 <!-- cor.test(brain_weight, fiber_count, method="kendall")  -->
<!-- 
The estimate for *Kendall's sample rank correlation coefficient* is

\[
    \hat{\tau} = 0.2727273
\]
 -->


## Solution to Problem 5

```{r}
cysticerci <- c(28.9, 32.8, 12.0, 9.9, 15.0, 38.0, 12.5, 36.5, 8.6, 26.8)
worms_reco <- c(1.0, 7.7, 7.3, 7.9, 1.1, 3.5, 18.9, 33.9, 28.6, 25.0)
```

The null hypothesis is that the mean weight of introduced cysticerci *has no correlation with* the mean weight of worms recovered. That is,

\[
    H_0: r_s < r_{s,\alpha}
\]

The alternative hypothesis is that the mean weight of introduced cysticerci is *positively correlated with* the mean weight of worms recovered. That is,

\[
    H_A: r_s \geq r_{s,\alpha}
\]

Otherwise, do not reject.

To test the null hypothesis against the alternative hypothesis, we will use the Spearman test, a distribution-free test for independence based on ranks.

<!-- # install.packages("SuppDists", dependencies=TRUE, repos='http://cran.us.r-project.org') -->

<!-- # qSpearman(0.01, r = 7) -->
```{r}
# this method of performing the test was given in the textbook
library(SuppDists)
qSpearman(p = 0.05, r = 10)
```

~~Since $r_{s,\alpha} = -0.5393939$, we will reject the null hypothesis only if $r_s \geq -0.5393939$.~~

~~Calculating $r_s$,~~

```{r}
cor(x = cysticerci, y = worms_reco, method = "spearman")
```

~~Since $r_s = -0.2$ and $r_{s,\alpha} = -0.5393939$, the statement $r_s \geq r_{s,\alpha}$ is *true*. Thus, we *reject* the null hypothesis. There is *sufficient* evidence that the mean weight of introduced cysticerci is *positively correlated with* the mean weight of worms recovered.~~

---

~~**NOTE**: At this point, I tried to use `cor.test()` with `method = "spearman"` but I got a different result than I expected, and I'm not sure why. Maybe I'm interpreting the output incorrectly?~~

```{r}
cor.test(x = cysticerci, y = worms_reco, method = "spearman", alternative = "greater")
```

The $p$-value is $0.72$, which is *not* significant at the $\alpha = 0.05$ level. There is *not enough* evidence that the mean weight of introduced cysticerci is *positively correlated with* the mean weight of worms recovered.

---

## Solution to Problem 6

```{r}
x = c(0, 5000, 10000, 15000, 20000, 25000, 30000, 100000)
y = c(0.924, 0.988, 0.992, 1.118, 1.133, 1.145, 1.157, 1.357)
```

The null hypothesis is that the mean weight of introduced cysticerci *has no correlation with* the mean weight of worms recovered. That is,

\[
    H_0: \beta = \beta_{0}
\]
\[
    H_0: \beta = 0
\]

The alternative hypothesis is that the mean weight of introduced cysticerci is *positively correlated with* the mean weight of worms recovered. That is,

\[
    H_A: \beta > \beta_{0}
\]
\[
    H_A: \beta > 0
\]

To test the null hypothesis against the alternative hypothesis, we will use the Theil test, a distribution-free test for the slope of the regression line.

```{r}
library(NSM3)
theil(x, y, alpha=0.05, beta.0=0, type = "u")
theil.fit = theil (x, 
  y,
  beta.0 = 0 ,
  slopes=TRUE,
  type = "u", 
  doplot = FALSE)
theil.fit
theil.output = theil(x, 
  y,
  beta.0 = 0,
  slopes=TRUE,
  type = "u", doplot = FALSE, alpha = .05)
c(theil.output$L, theil.output$U)
```

**TODO**: Interpret these results correctly.

## Solution to Problem 7

```{r}
height = c(42.8, 63.5, 37.5, 39.5, 45.5, 38.5, 43.0, 22.5, 37.0, 23.5, 33.0, 58.0); length(height)
weight = c(40.0, 93.5, 35.5, 30.0, 52.0, 17.0, 38.5, 8.5, 33.0, 9.5, 21.0, 79.0); length(weight)
heart_catheter_length = c(37.0, 49.5, 34.5, 36.0, 43.0, 28.0, 37.0, 20.0, 33.5, 30.5, 38.5, 47.0); length(heart_catheter_length)

model1 <- lm (heart_catheter_length ~ height + weight)
summary(model1)
cor.test(x = height, y = weight, method = "pearson")
confint(model1, conf.level = 0.95)

library(Rfit)
rfit(heart_catheter_length ~ height + weight)
rankfit_object <- rfit(heart_catheter_length ~ height + weight)
summary(rankfit_object)

r.01 <- rfit(heart_catheter_length ~ height)
f.01 <- rfit(heart_catheter_length ~ height + weight)
drop.test(f.01, r.01)
```